JavaScript as an Embedded DSL

### 2012-03-21 papers/ecoop2012/article.tex
contributions from rebuttal

We propose to embed JavaScript as a DSL in a host language. Through LMS, we tackle the challenges outlined above with minimal effort, as most of the work is off-loaded to the host language. In particular, we make the following contributions:
\begin{itemize}
\item Our DSL is statically typed through the host language, yet supports gradual typing notably for incorporating external JavaScript libraries and APIs (section~\ref{sec:typing}).
\item In addition to generating JavaScript code, our DSL can be executed directly in the host language, allowing code to be shared between client and server (section~\ref{sec:trivial-embedding}).
\item We use advanced object-oriented techniques to achieve modularity in our DSL: each language primitive and API is defined in a separate module (section~\ref{sec:modularity}).
\item Our DSL supports typed object literals and class-based objects. The translations to JavaScript are lightweight and intuitive: the object literals translate to JSON-like object literals and the class-based objects to JavaScript constructor-based objects (section~\ref{sec:reification}).
\item With minimal effort, we exploit the selective CPS transform already existing in the host language to provide a compelling abstraction over asynchronous callback-driven programming in our DSL (section~\ref{sec:cps}). This case-study demonstrates the fruitfulness of re-using host language features to enhance our embedded DSL.
\end{itemize}

@@

\subsection{Casting and Optional Runtime Type Checks}
The need for casting arises in a few contexts. One of them is the
boundary between typed and untyped portions of a
program~\cite{wadler-blame,blame-for-all}. Passing a value from an
untyped portion to a typed one usually requires a cast. Another
situation where casts are needed is interaction with external
services. For example, to process data from an external service such
as Twitter, we cast it to its expected type:
\begin{lstlisting}
type TwitterResponse = Array[JSLiteral {val text: String}]
def fetchTweets(username: Rep[String]) = {
 val raw = ajax.get { ... }
 raw.as[TwitterResponse]
}
\end{lstlisting}
A more complete example is provided in section~\ref{sec:cps}. In this
situation, it is useful to generate runtime checks either as an aid
during development and debugging time or as a security mechanism that
validates data coming from an external source. In the example above,
if runtime checks are enabled, by failing early, we obtain a guarantee
that all data returned from \code{fetchTweets} conforms to type
\code{TwitterResponse} which means that any later access to the
\code{text} field of any element of the data array will never fail, and
always return a string.

@@

\subsection{Scala Dynamic}
Since our DSL compiles to JavaScript, which is dynamically typed, it
is appealing to allow expressions and APIs in our DSL to also,
selectively, be dynamically typed. This is especially useful for
rapid-prototyping.

@@

\subsection{From Dynamic to Static}

@@

\subsection{Classes}
Our DSL also supports reified classes. For convenience, these are
defined as traits, use the \code{repProxy} mechanism underlying typed
APIs and are also implemented using reflection. The translation to
JavaScript, based on constructors and prototypes, is
straightforward. Through the trivial embedding, classes implemented in
our DSL can be used on both the server and client sides.

>>>>

We propose to embed JavaScript as a DSL in a host language. Through LMS, we tackle the challenges outlined above with minimal effort, as most of the work is off-loaded to the host language. In particular, we make the following contributions:
\begin{itemize}
\item Our DSL is statically typed through the host language, yet supports gradual typing notably for incorporating external JavaScript libraries and APIs (section~\ref{sec:typing}).
\item In addition to generating JavaScript code, our DSL can be executed directly in the host language, allowing code to be shared between client and server (section~\ref{sec:trivial-embedding}).
\item We use advanced object-oriented techniques to achieve modularity in our DSL: each language primitive and API is defined in a separate module (section~\ref{sec:modularity}).
\item Our DSL supports typed object literals and class-based objects. The translations to JavaScript are lightweight and intuitive: the object literals translate to JSON-like object literals and the class-based objects to JavaScript constructor-based objects (section~\ref{sec:reification}).
\item With minimal effort, we exploit the selective CPS transform already existing in the host language to provide a compelling abstraction over asynchronous callback-driven programming in our DSL (section~\ref{sec:cps}). This case-study demonstrates the fruitfulness of re-using host language features to enhance our embedded DSL. A novel insight is that CPS transforming a program generator allows it to generate code that is in CPS.
\item The present work extends both the range of virtualizable language features and the LMS framework, which is beneficial to future DSL efforts taking the JavaScript work as a case study. Previous LMS embeddings had to define each operation explicitly (like in section~\ref{sec:defineDslComponent}). This paper contributes lifting of whole traits or classes (through the \code{repProxy} mechanism described and used in sections~\ref{sec:typedApis}\&~\ref{sec:reified-classes}), untyped or optionally typed operations (sections~\ref{sec:casting},~\ref{sec:dynamic}\&~\ref{sec:dynamic2static}), and typed object literals (section~\ref{sec:object-literals}) including the necessary Scala-Virtualized support.
\end{itemize}

@@

\subsection{Casting and Optional Runtime Type Checks}\label{sec:casting}
The need for casting arises in a few contexts. One of them is the
boundary between typed and untyped portions of a
program~\cite{wadler-blame,blame-for-all}. Passing a value from an
untyped portion to a typed one usually requires a cast. Another
situation where casts are needed is interaction with external
services. For example, to process data from an external service such
as Twitter, we cast it to its expected type:
\begin{lstlisting}
type TwitterResponse = Array[JSLiteral {val text: String}]
def fetchTweets(username: Rep[String]) = {
 val raw = ajax.get { ... }
 raw.as[TwitterResponse]
}
\end{lstlisting}
A more complete example is provided in section~\ref{sec:cps}. In this
situation, it is useful to generate runtime checks either as an aid
during development and debugging time or as a security mechanism that
validates data coming from an external source. In the example above,
if runtime checks are enabled, by failing early, we obtain a guarantee
that all data returned from \code{fetchTweets} conforms to type
\code{TwitterResponse} which means that any later access to the
\code{text} field of any element of the data array will never fail, and
always return a string.

@@

\subsection{Scala Dynamic}\label{sec:dynamic}
Since our DSL compiles to JavaScript, which is dynamically typed, it
is appealing to allow expressions and APIs in our DSL to also,
selectively, be dynamically typed. This is especially useful for
rapid-prototyping.

@@

\subsection{From Dynamic to Static}\label{sec:dynamic2static}

@@

\subsection{Classes}\label{sec:reified-classes}
Our DSL also supports reified classes. For convenience, these are
defined as traits, use the \code{repProxy} mechanism underlying typed
APIs and are also implemented using reflection. The translation to
JavaScript, based on constructors and prototypes, is
straightforward. Through the trivial embedding, classes implemented in
our DSL can be used on both the server and client sides.
